Exception Handling Bug Hazards in Android

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes

In this section each piece information is then analyzed in detail to 
check whether they can reveal bug harzards on the exception handling code -
related to (i) specific violations of the best practices presented in Section~\ref{sec:best},
or (ii) the general use of exception handling to suport robust development.

@@

In this section, we present our findings per research question. 
We also include a selection of quotes from the exploratory survey 
respondents. Each respondent has an identification which can be traced in our database.

@@

@@

@@

\greybox{They can happen pretty much anywhere.	 The Android Fragment system comes to mind in this case. Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for. This might have been avoidable under a different structure.}

\greybox{The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did.}

@@

\greybox{Honestly, there are a lot of very unskilled Java programmers out there writing Android apps. When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.}

%\greybox{This is nothing specific to Android, it's just how Java works.}

@@

\greybox{Determine why the object was null and attempt to fix this situation. In the case of external API calls which return null, then check for null (the quick-and-dirty way). For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null.}

\greybox{Probably the most common exception, as such the reasons and fixes are super varying. But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message. Which will communicate useful information without the confusion an NPE would normally cause.}

@@

@@

\greybox{I am guessing this is because the exception is thrown from native code (ie, C++ code in the JVM) where Java correct-ness semantics rules do not always apply.}

@@

\greybox{Any uncatched exception will crash the app. There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT}

\greybox{The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called.}

@@

\greybox{Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events. It may be hard to ensure that all this cases are properly handled. }

\greybox{You can wrap the main function in java in try/catch in android you can't. }

>>>>

In this section each piece information is analyzed in detail to 
check whether they can reveal bug harzards on the exception handling code -
related to (i) specific violations of the best practices presented in Section~\ref{sec:best},
or (ii) the general use of exception handling to suport robust development.

@@

In this section, we present our findings per research question. 
To illustrate the different aspects of each
finding, we provide a selection of quotes from the exploratory
survey. 
To enable traceability, each respondent has an identification
 which can be traced in our database using the following convention: 
D\#. For instance, D1 corresponds to the answer provided
 by developer 1.

@@

``Don't throw RuntimeException (in most cases), don't catch OutOfMemoryException, etc." [D42]

``If a method signature declares that it throws an exception, it means that the caller should handle the exception. [D29]

@@

\greybox{``They [NullPointerExceptions] can happen pretty much anywhere. The Android Fragment system comes to mind in this case. Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for. This might have been avoidable under a different structure."[D56]}

@@

%``android"

@@

\greybox{``Honestly, there are a lot of very unskilled Java programmers out there writing Android apps. When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.
This is nothing specific to Android, it's just how Java works."[D42]}

\greybox{``The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did." [D58]}

\greybox{``Android destroys and recreates itself all of the time ( especially during screen rotation ) and if you do not handle that it will crash on you every time. with the complexity of an activity with a fragment that has fragments and each of those fragments has custom objects and variables that need to be either retained ( so saved and put back ) or recreated such as views it can get complex if you don't have an understanding of how the android life cycle works." [D43]}

@@

\greybox{``Determine why the object was null and attempt to fix this situation. In the case of external API calls which return null, then check for null (the quick-and-dirty way). For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null." [D42]}

\greybox{``Probably the most common exception, as such the reasons and fixes are super varying. But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message. Which will communicate useful information without the confusion an NPE would normally cause."[D48]}

@@

``RuntimeException-s are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]

``this treats all exceptions as critical failure, even though they might not all be unrecoverable" [D70]

``Its very important in android that we gracefully handle an exception. Try not to crash an application."[41]

``Rethrowing a checked exception as an unchecked exception is the worst choice; it throws away any value of checked exceptions, elevates the exception out of almost all handling, and does absolutely nothing helpful in the process. Further, catching `Exception` suggests that the author doesn't know what exceptions might occur and wrongfully assumed that this solution is somehow safer than doing nothing."[D19]

``Giving the user meaningful feed back is important, so the more specific you can be about what went wrong, the better."[D66]

``because it is checked exception, you should expect it to happen rather than just crash"[D70]

@@

\greybox{``I am guessing this is because the exception is thrown from native code (ie, C++ code in the JVM) where Java correct-ness semantics rules do not always apply." [D42]}

@@

\greybox{``Any uncatched exception will crash the app. There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT"[D32]}

\greybox{``The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called."[D70]}

@@

\greybox{``Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events. It may be hard to ensure that all this cases are properly handled." [D46]}

\greybox{``You can wrap the main function in java in try/catch in android you can't..."[D12]}
